Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity\vspace{-7ex}

### 2016-08-06 papers/kmeans/kmeans.tex
ting references; holtzmann comment

We have demonstrated for the first time that, in a large-scale
spectroscopic survey, stars that appear similar in
chemical-abundance properties are often associated physically in
clusters or other structures.
The reverse is well established---that is, that stars that are
associated physically in clusters are similar in chemical-abundance
properties.
That is, although they often show chemical diversity, that chemical
diversity is small, and follows well-defined trends or a subspace in
abundance space (for example, \citealt{gratton, meszaros, bovy},
and references within those).
However, it was not known---prior to this work---whether abundance-space information would be
informative enough, or \emph{measurable} precisely enough to find this
structure in surveys dominated by a large mix of stars from different
origins and ages (and there was good reason to be pessimistic;
\citealt{ting}).
We have demonstrated with a straightforward (and in no way optimal)
clustering in chemical-abundance space that it is easy to find co-eval
stellar structures.
Although this is very strong evidence that chemical tagging will work,
we have not performed, in any sense, the canonical procedure of
looking for features that are consistent with being delta functions in
chemical-abundance space (convolved with a measurement noise
distribution); we have looked only for over-densities, not delta
functions.

@@

%HOGG: THIS PARAGRAPH IS ALL WRONG!
There are scant abundance measurements reported for Palomar 5, for which
most stars in the cluster in \figurename~\ref{fig:Pal5} belong. We find a mean
metallicity of [Fe/H]$ = -1.3$, which lies between the only
photometric \citep{Geisler_1997} and spectroscopic \citep{Smith_1985,Smith_2002} abundance
measurements for this cluster.  The anti-correlation in C--N and Na--O abundance
have been seen elsewhere \citep[][respectively]{Smith_1985,Smith_2002}, although
to our knowledge the Mg--Al anti-correlation has not previously been shown for this cluster.
Indeed, this work contains the largest number of abundances for Palomar 5,
permitting a thorough search of tidally disrupted members \citep[for example,][]{Kuzma_2015} 
that have retained the unique cluster abundance signature.

>>>>

We have demonstrated for the first time that, in a large-scale
spectroscopic survey, stars that appear similar in
chemical-abundance properties are often associated physically in
clusters or other structures.
The reverse is well established---that is, that stars that are
associated physically in clusters are similar in chemical-abundance
properties.
That is, although they often show chemical diversity, that chemical
diversity is small, and follows well-defined trends or a subspace in
abundance space (for example, \citealt{gratton, meszaros, bovy},
and references within those).
However, it was not known---prior to this work---whether abundance-space information would be
informative enough, or \emph{measurable} precisely enough to find this
structure in surveys dominated by a large mix of stars from different
origins and ages.
There was good reason to be pessimistic, but plausibly the increased
precision obtained here by \thecannon\ made this possible; even tiny improvements
in individual abundances can enormously increase contrast in the high-dimensional
abundance space \citep{ting}.
We have demonstrated with a straightforward (and in no way optimal)
clustering in chemical-abundance space that it is easy to find co-eval
stellar structures.
Although this is very strong evidence that chemical tagging will work,
we have not performed, in any sense, the canonical procedure of
looking for features that are consistent with being delta functions in
chemical-abundance space (convolved with a measurement noise
distribution); we have looked only for over-densities, not delta
functions.

One clear result here is that it is easier for the clustering algorithm
to find structures at low metallicity than at high metallicity:
Some of our low-metallicity features are very compact in phase space,
while none of our higher metallicity features are.
This probably relates to the much higher background
(in abundance space) of unclustered stars at higher metallicities than
at low metallicities; in general it is easier to find abundance-space features
at the edges of the distribution than in the center \citep{ting}.

@@

% HOGG: THIS PARAGRAPH IS ALL WRONG!
% There are scant abundance measurements reported for Palomar 5, for which
% most stars in the cluster in \figurename~\ref{fig:Pal5} belong. We find a mean
% metallicity of [Fe/H]$ = -1.3$, which lies between the only
% photometric \citep{Geisler_1997} and spectroscopic \citep{Smith_1985,Smith_2002} abundance
% measurements for this cluster.  The anti-correlation in C--N and Na--O abundance
% have been seen elsewhere \citep[][respectively]{Smith_1985,Smith_2002}, although
% to our knowledge the Mg--Al anti-correlation has not previously been shown for this cluster.
% Indeed, this work contains the largest number of abundances for Palomar 5,
% permitting a thorough search of tidally disrupted members \citep[for example,][]{Kuzma_2015} 
% that have retained the unique cluster abundance signature.

[HOGG, CASEY: INSERT PARAGRAPH ABOUT M5 HERE]
